---
title: "Motor insurance analysis_2"
author: "Omar Sou'b"
date: "`r Sys.Date()`"
output:
 rmdformats::readthedown 
 
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# introduction

this analysis is conducted by using R-programming language as an analysis tool in the insurance industry, and specially for motor insurance .

the data set used called (datacar) and can be found on the (insurancedata) package on r.

**\>we have also used packages such as:**

-   tidyvarse.

-   tidyr.

-   janitor.

-   knitr.

-   insurancerating , and others.

**\>what we are going to do ?**

-   looking at the data summary and description on order to understand the shape of the data.

-   calculating the claims cost and group it by factors.

-   calculating the claims count and group it by factors.

-   calculating the total exposure and group it by factors.

-   calculating the frequency cost and group it by factors.

-   calculating the average severity cost and group it by factors.

-   calculating the pure premium cost and group it by factors.

-   calculating the rating factor for each.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(janitor)
library(skimr)
library(ggplot2)
library(mosaic)
library(patchwork)
library(knitr)  # in order to use kable()
library(GGally)
library(broom)
library(MASS)

library(flexplot)
library(insuranceData)
library(insurancerating)

```

**----------------------------------------------**

# Part One: data set view,summary-descriptive analysis



```{r message=FALSE, warning=FALSE, include=FALSE}
data("dataCar")

dataCar<-dataCar[,-11]
```

------------------------------------------------------------------------

## data set dimensions

lets have a look at the data set dimension :

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar %>% dim() 
```

------------------------------------------------------------------------

the data set has 10 variables and and 67856 rows , the 1st 5 rows are as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar %>% head() %>%kable()
```

------------------------------------------------------------------------

## data set variables types

lets have a look at the variables types:

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar %>% glimpse() 

```

------------------------------------------------------------------------

now we will make some data transformation,will do the below :

1.  convert agecat & veh_age types into factor

2.  create a new variable called: veh_value_group

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar$agecat<-as.factor(dataCar$agecat)
dataCar$veh_age<-as.factor(dataCar$veh_age)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

dataCar<-dataCar %>% mutate(veh_value_group=as.factor(ifelse(veh_value >=0 & veh_value <=10,{"group 1"}, 
                   ifelse(veh_value >10 & veh_value <=20,{"group 2"},
                   ifelse(veh_value >20 & veh_value <=30,{"group 3"},
                   {"group 4"}))))) 
```

------------------------------------------------------------------------

lets look again to the variables types :

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar %>% glimpse() 
```

------------------------------------------------------------------------

## data set summary and distribution

lets have the data summary and the variables distribution :

```{r echo=FALSE, message=FALSE, warning=FALSE}
dataCar %>% summary() %>% kable()
```

------------------------------------------------------------------------

### Numerical variables

```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
par(mfrow=c(1,3))
boxplot(dataCar$veh_value,main="veh_vale distribution")
boxplot(dataCar$exposure,main="exposure distribution")
boxplot(dataCar$claimcst0,main="claimcst0 distribution")
par(mfrow=c(1,1))

```

-   the vehicle value range is between 0 and 34,560 with 1,777 on average

-   the exposure range is between 0 and 1 with 0.44 on average

-   the claims cost range is between 0 and 55,922.1 with 137,3 on average

------------------------------------------------------------------------

### categotical variables

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
(dataCar %>% ggplot(aes(clm))+geom_bar(aes(fill=clm,color=clm),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100),"%"))),stat='count',nudge_x=-0.001,
                                                        nudge_y=4000)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=1000)+
    dataCar %>% ggplot(aes(gender))+geom_bar(aes(fill=gender,color=gender),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100),"%"))),stat='count',nudge_x=-0.001,
                                                        nudge_y=3000)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=1000)+
    dataCar %>% ggplot(aes(veh_age))+geom_bar(aes(fill=veh_age,color=veh_age),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100),"%"))),stat='count',nudge_x=-0.001,
                                                        nudge_y=1800)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=1000))
```

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=TRUE}

   (dataCar %>% ggplot(aes(area))+geom_bar(aes(fill=area,color=area),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100),"%"))),stat='count',nudge_x=-0.002,
                                                        nudge_y=1300,size=3.5)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=500)+
    dataCar %>% ggplot(aes(agecat))+geom_bar(aes(fill=agecat,color=agecat),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100),"%"))),stat='count',nudge_x=-0.002,
                                                        nudge_y=1300,size=3.5)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=500)+
    dataCar %>% ggplot(aes(veh_value_group))+geom_bar(aes(fill=veh_value_group,color=veh_value_group),show.legend = FALSE)+geom_text(aes(label=after_stat(paste(round(count / sum(count) *100,2),"%"))),stat='count',nudge_x=-0.002,
                                                        nudge_y=3500,size=3.5)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=1000))
```

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}


dataCar %>% ggplot(aes(veh_body))+geom_bar(aes(fill=veh_body,color=veh_body),show.legend = FALSE)+ geom_text(aes(label=after_stat(paste(round(count / sum(count) *100,2),"%"))),stat='count',nudge_x=-0.001,
                                                        nudge_y=1500)+ geom_text(aes(label=after_stat(count)),stat='count',nudge_x=-0.001,nudge_y=500)



```

-   7% of the vehicles in this data had claims .

-   43% of the data samples are for males and 57% are for females.

-   more than 50% of our vehicles are old (3 & 4 ages) and only 18% are brand new vehicles.

-   area C has the highest number of customers with 30 % , and area F has the lowest 5%.

-   more than 45% of the customers are located in the age group 3 & 4.

-   more than 99% of our vehicles values are within group 1.

-   sedan vehicles has the highest percentage with 32% ,and (Bus,CONVT,MCARA,RDSTR) has the lowest with less than 1%.

**----------------------------------------------**

# Part Two: insurance risk measurements by factors.

as the insurance risk level is measured by the frequency and the severity; we will calculate the the same and group it by factors.

## gender factor

```{r echo=FALSE, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = gender, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% kable()
```

> we can see that :

-   females number of claims are more than males.

-   females claims cost are more than males but; the average severity is less ,and this is due to the larger number of claims(more than males)

-   both females and males almost have the same frequency.

-   female risk premium is lesser than males ,and this due to the higher exposure females have

the below chart visualize all together for more clear understanding:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
univariate(dataCar, x = gender, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% autoplot(background = FALSE)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
dataCar<-dataCar %>% mutate(across(c(veh_body,veh_age,gender,area,agecat),as.factor)) %>% 
  mutate(across(c(veh_body,veh_age,gender,area,agecat),~biggest_reference(.,exposure))) 

numclaims_glm_gender<-glm(data = dataCar,formula = numclaims~gender,family = poisson,offset = log(exposure))
```

------------------------------------------------------------------------

## veh_age factor

```{r echo=FALSE, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = veh_age, severity = claimcst0, nclaims = numclaims,
           exposure = exposure)  %>% kable()
```

> we can see that :

-   the claims cost for the vehicle age 1 is the lowest while for vehicle age 3 is the highest.

-   vehicle age 1 has the lowest number on claims while the vehicle age 3 has the highest.

-   the exposure for the vehicle age 1 is the lowest while for vehicle age 3 is the highest.

-   the frequency for the vehicle age 4 is the lowest while for vehicle age 2 is the highest.

-   the average_severity for the vehicle age 1 is the lowest while for vehicle age 4 is the highest.

-   the risk premium for the vehicle age 3 & 4 is the lowest and almost the same, while its the highest for vehicle age 2

the below chart visualize all together for more clear understating :

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}

univariate(dataCar, x = veh_age, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% autoplot(background = FALSE)

```

```{r message=FALSE, warning=FALSE, include=FALSE}

numclaims_glm_veh_age<-glm(data = dataCar,formula = numclaims~veh_age,family = poisson,offset = log(exposure))

```

------------------------------------------------------------------------

## veh_body

```{r echo=FALSE, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = veh_body, severity = claimcst0, nclaims = numclaims,
           exposure = exposure)%>% kable()
```

> We can see that:

-   sedan,Hback,and Stnwg have the highest claims cost (), while RDSTR has the lowest claims cost.

-   RDSTR ,convt,Bus,MCARA,COUPE & PANVN have less than 100 claims, SEDAN,STNWG & HBACK have more than 1000 claims ,other have claims between 130 and 276.

-   BUS has the highest frequency with 38.3% , (RDSTR,MCARA, and COUPE have frequincies (27.7%,25.3% and 23.5%), other groups have frequencies less than 18%.

-   RDSTR has the lowest average severity , (MIBUS,COUPE and TRUCK ) have the highest.

-   the risk premium is the less than 200 for RDSTR and MCARA, the more tgan 500 for BUS and COUPE.

the below chart visualize all together for more clear understanding:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = veh_body, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% autoplot(background = FALSE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
numclaims_glm_veh_body<-glm(data = dataCar,formula = numclaims~veh_body,family = poisson,offset = log(exposure))
```

------------------------------------------------------------------------

## area

```{r echo=FALSE, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = area, severity = claimcst0, nclaims = numclaims,
           exposure = exposure)%>% kable()
```

> we can see that :

-   the lowest claims cost is for area F , and the highest cost is for area C.

-   area F has the lowest number of claims , while area C has the highest .

-   area D has the lowest frequency , while area F has the highest.

-   area D has the lowest average frequency ,while area F has the highest.

-   area D has the lowest risk premium , while area F has the highest.

the below chart visualize all together for more clear understanding:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
####
univariate(dataCar, x = area, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% autoplot(background = FALSE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
numclaims_glm_area<-glm(data = dataCar,formula = numclaims~area,family = poisson,offset = log(exposure))
```

------------------------------------------------------------------------

## age category

```{r echo=FALSE, message=FALSE, warning=FALSE}
univariate(dataCar, x = agecat, severity = claimcst0, nclaims = numclaims,
           exposure = exposure)%>% kable()
```

> we can see that :

-   age category 6 has the lowest claims cost , while age category 4 has the highest

-   age category 6 has the lowest number of claims , while age category 4 & 3 has the highest (1185,1189) .

-   age category 6 & 5 has the lowest frequency with almost(12.5%) , while age category 1 has the highest with 20%.

-   age category 5 has the lowest average frequency ,while age category 1 has the highest.

-   age category 5 has the lowest risk premium , while age category 1 has the highest.

the below chart visualize all together for more clear understanding:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
univariate(dataCar, x = agecat, severity = claimcst0, nclaims = numclaims,
           exposure = exposure) %>% autoplot(background = FALSE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
numclaims_glm_agecat<-glm(data = dataCar,formula = numclaims~agecat,family = poisson,offset = log(exposure))



```

------------------------------------------------------------------------

now, lets create a new model that contains all factor variables and see how much it differs from the previous models and which one is the best.

```{r message=FALSE, warning=FALSE, include=FALSE}
numclaims_glm_all<-glm(formula = numclaims~veh_body+veh_age+gender+area+agecat,family = poisson,offset = log(exposure),data = dataCar)

```

```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
model.comparison(numclaims_glm_all,numclaims_glm_agecat) 
model.comparison(numclaims_glm_all,numclaims_glm_area)
model.comparison(numclaims_glm_all,numclaims_glm_gender)
model.comparison(numclaims_glm_all,numclaims_glm_veh_age)
model.comparison(numclaims_glm_all,numclaims_glm_veh_body)
```

> from our results we can say that the numclaims_glm_all model is best .

------------------------------------------------------------------------

lets apply the anova function on the model so we can well know how much the variables are important to the model

```{r echo=FALSE, message=FALSE, warning=FALSE}
anova(numclaims_glm_all,test = "Chisq")
```

> The drop in deviance allows to test whether the model terms are significant :

-   adding veh_body to the NULL model causes a drop in deviance of 37.563 corresponding to 67843 degree of freedom and resulting residual deviance of 25469

-   adding veh_age to the NULL model causes a drop in deviance of 33.543 corresponding to 67840 degree of freedom and resulting residual deviance of 25436

-   adding gender to the NULL model causes a drop in deviance of 1.595 corresponding to 67839 degree of freedom and resulting residual deviance of 25434

-   adding area to the NULL model causes a drop in deviance of 14.525 corresponding to 67834 degree of freedom and resulting residual deviance of 25420

-   adding agecat to the NULL model causes a drop in deviance of 86.074 corresponding to 68729 degree of freedom and resulting residual deviance of 25334

so the gender variable has the least impact on the model

**-----------------------------**

# Part Three: claims count response to the predictors variables

the below chart express visually how number of claims will act (decrease on increase)-along with the lower and upper confident intervals-, when any of the factors increase by one unit:

```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=TRUE}
numclaims_glm_all %>% coefplot::coefplot(title="all factors coefficient",)
```

> if any factor increases by one unit ;the number of claims will increase or decrease by the coefficient value for the same factor.

**-----------------------------**

# Part four: rating factors

now;lets look at the rating factor for the new model (numclaims_glm_all) which is the best performed model:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
rating_factors(model_data = dataCar,numclaims_glm_all,exposure = exposure) %>% kable()
```

on graph:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
rating_factors(model_data = dataCar,numclaims_glm_all,exposure = exposure) %>% autoplot()
```

**-----------------------------**

# Conclusion: 

## preferred segments

-   area : D.

-   agecat : 5,6

-   veh_body : UTE,CONVT.

## the high risk segments

-   veh_body : BUS,COUPE,MACARA,RDSTS

------------------------------------------------------------------------
