---
title: "Claims Reserve"
author: "Omar"
date: "2022/05/02"
output:
 html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false

---


# introduction 

the purpose of this project is to predict the best claims reservation (incurred claims+IBNR)
noting that the incurred claims means the claims incurred and have been reported to the insurance company,while 
the IBNR is the claims incurred but not reported (pure IBNR , or incurred but not enough reported IBNER due to claims developments)


we will be using - autoPaid - data set in triangle format/type which can be found on chainladder package,noting that we will be using chainladder package from R analysis software especially for achieving our project.



```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(janitor)
library(skimr)
library(ggplot2)
library(knitr)  # in order to use kable()
library(insuranceData)
library(insurancerating)
library(report)
library(flextable)
library(ChainLadder)

```


# Part One : The data set.

1. We can look at the data set in triangle format as below :

```{r echo=FALSE, message=FALSE, warning=FALSE}
####data
autoPaid<-auto$PersonalAutoPaid
autoPaid<-as.triangle(autoPaid,
                      origin="origin",
                      dev="dev",
                      value="value")
autoPaid 
```

> the above data set show the claims incurred development by each development year for each origin year.
in instance the claims incurred for the 1st year were 101125 , then it became 209921 in the 2nd development year due to claims developments which include IBNR and IBNER , and the same apply for all origin and dev years.

**note that the 10nth development year for the 1st origin year has the ultimate reserve which include incurred and IBNR**

  + on graph:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot(autoPaid,lattice=TRUE)
```


---------------------------------------------

---------------------------------------------

2. lets look at the development years cumulative claims each origin year:

```{r echo=FALSE, message=FALSE, warning=FALSE}
####data increment by development year
incr_autoPaid<-incr2cum(autoPaid)
incr_autoPaid
```

  + on graph:
```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot(incr_autoPaid,lattice=TRUE)
```

---------------------------------------------

---------------------------------------------

3. lets look on the difference  between the development years claims for each origin year:

```{r echo=FALSE, message=FALSE, warning=FALSE}
####data cumulative by develpment year
cum_autoPaid<-cum2incr(autoPaid)
cum_autoPaid 
```

---------------------------------------------

---------------------------------------------

4. lets look at the percent of changing in the development years ***claims*** each origin year.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ate_autoPaid<-ata(autoPaid)
ate_autoPaid 
```
> we can see that for the 1st origin year, the differnce percent between the 1st &2nd development years is 2.076 and between the 2nd & 3rd development years is 1.270 and so on, this means, that the iccured claims increased by 2.076 on the 2nd development year & in creased by 1.270 on the 3rd development year

---------------------------------------------

# Part two : mackchainladder

now will apply the ***mackchainladderfuntion*** model to forecasts future claims developments based on a historical  claims development triangle and estimates the standard error around those.


```{r echo=FALSE, message=FALSE, warning=FALSE}
####mackchainladder
mack_autoPaid<-MackChainLadder(autoPaid,est.sigma = "Mack")
mack_autoPaid
```

>we can see that:

+ the Ultimate for the 1st origin year is 353,584 & for the 2nd year origin 350,752 and so on.

+ the IBNR for the 2nd year is 229 & for the 3rd year 1830 and so on.

+ the total of latest claims is 3,290,539.00 & the total ultimate claims is 3,914,785.82 with .87 Dev to date in total.

+ the total IBNR is 624,246.82 with mack standard error of 0.05 in total (30,358.21 in numbers)





on graph :

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
mack_autoPaid %>% plot()
mack_autoPaid %>% plot(lattice=TRUE)

```


---------------------------------------------

---------------------------------------------

+ to see the standard error in triangles :

```{r echo=FALSE, message=FALSE, warning=FALSE}
mack_autoPaid$Mack.S.E  ## age to age factors and standard errors of the chain-ladder age-to-age factors f (assumption CL1)
```
> we can see that the 2nd origin year will have almost 998 S.E at the 10th Dev. year, & the 3rd origin year will have almost 1333 S.E at the 9th Dev. year ,and almost 1713 at the 10th Dev. yearm ans so on.

---------------------------------------------

---------------------------------------------

+ to see the full triangle :

```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
mack_autoPaid$FullTriangle
```




---------------------------------------------

# Part Three : data distrinbution


Now, lets look the the distribution of the data using ***bootchainladder*** function :


```{r echo=FALSE, message=FALSE, warning=FALSE}
####BootChainLadder
BootChainLadder_autoPaid <-BootChainLadder(autoPaid,R=999,process.distr = c("gamma","od.pois"))
BootChainLadder_autoPaid 
```

>we can see that:

+ the mean ultimate inccured claims is : 3,915,580

+ the mean IBNR  is : 625,041 with standard error of : 30,442

+ the IBNR at the 75% quantile is:645,833 & and at 95% is :675,558

+ this means that the IBNR is almost normally distributed .


on graph:

```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
BootChainLadder_autoPaid %>% plot()
```

---------------------------------------------

# Part Four : reserve using GLM model


+ some insurance companies use the generalized linear model for loss reserving .

```{r echo=FALSE, message=FALSE, warning=FALSE}
####glmreserve
glm_Reserve_autoPaid <-glmReserve(autoPaid)  
glm_Reserve_autoPaid


```

+ We can also extract the underlying GLM model :

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(glm_Reserve_autoPaid,type="model")
```

---------------------------------------------

# Part Four : Paid-incurred chain model



this model uses both : the claims payments and incurred losses(the number of incuured losses) to get a  unified ultimate loss prediction.

will use the same paid claims data used earlier, and the incurred claims data set we will use is the -autoIncurred- which can be found on chainladder package

lets have a look at both data sets:

+ autoPaid data set:

```{r echo=FALSE, message=FALSE, warning=FALSE}
autoPaid

```


+ autoIncurred data set:(no. the incurred losses)

```{r echo=FALSE, message=FALSE, warning=FALSE}
autoIncurred<-auto$PersonalAutoIncurred
autoIncurred<-as.triangle(autoIncurred,
                          origin="origin",
                          dev="dev",
                          value="value")
autoIncurred
```


+ now lets look at the model outputs:

```{r echo=FALSE, message=FALSE, warning=FALSE}
Paid_icurred<-PaidIncurredChain(autoPaid,autoIncurred) 

Paid_icurred

```

> we can see that

+ the  ultimate loss  will be 3489226

+ the  total reserve  will be 552271.4

+ the  total S.E  will be 21612.81

--------------------------------------------

# Part Four : One year claims development


this model means a short-term view and assessments of the one-year changes of the claims predictions when one updates the available information at the end of each accounting year

we will use the functions CDR : -one year claim development result -

```{r echo=FALSE, message=FALSE, warning=FALSE}
#CDR(mack_autoPaid)   # can be used with MackChainLadder & BootChainLadder
#CDR(BootChainLadder_autoPaid) 

CDR(mack_autoPaid,dev = "all") #to review the full claims development picture set the argument 
```

---------------------------------------------

# Part Five : glmreserve


many issuance uses for loss reserving:

+ why? when over-dispersed Poisson model is used, it reproduces the estimates from chain-ladder;

+ it provides a more coherent modelling framework than the Mack method;

+ all the relevant established statistical theory can be directly applied to perform hypothesis testing and diagnostic checking;

```{r echo=FALSE, message=FALSE, warning=FALSE}
glm_Reserve_autoPaid <-glmReserve(autoPaid)  
glm_Reserve_autoPaid 
```

to extract the underlying GLM model:

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(glm_Reserve_autoPaid,type="model") 
```


---------------------------------------------

# Part Five : Clark’s methods


>Clark’s methods work on incremental losses. His likelihood function is based on the assumption that incremental losses follow an over-dispersed Poisson (ODP) process.

+ LDF : Loss Development factor/Format Clark's method/ Losses develop according to a theoretical growth curve.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ClarkLDF(autoPaid) 
ClarkLDF(autoPaid,maxage = 20) 

```

on graph:

```{r echo=FALSE, message=FALSE, warning=FALSE}
ClarkLDF(autoPaid,maxage = 20) %>% plot()
```


***but;*** using The Weibull growth curve tends to be faster developing than the log-logistic

```{r echo=FALSE, message=FALSE, warning=FALSE}
ClarkLDF(autoPaid,G="weibull") 
```

on grapgh:

```{r echo=FALSE, message=FALSE, warning=FALSE}
ClarkLDF(autoPaid,G="weibull") %>% plot()
```


---------------------------------------------

+ LDF Clark’s Cap Cod method/expected ultimate losses in each origin year are the product of earned premium that year and a theoretical loss ratio.

```{r echo=FALSE, message=FALSE, warning=FALSE}
####ClarkCapeCod
ClarkCapeCod(autoPaid,Premium = 1000000,G="weibull")
```

on grapgh:

```{r echo=FALSE, message=FALSE, warning=FALSE}
ClarkCapeCod(autoPaid,Premium = 1000000,G="weibull") %>% plot()
```

